---
title: "Tidy Data"
author: "Ziwei Zhao"
date: "9/30/2021"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
```

## `pivot_longer`

```{r}
pulse_df <- 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()
pulse_df
```
Problem: the BDI score is spread across four columns, which correspond to four observation times. We can fix this problem using `pivot_longer`:

```{r}
pulse_tidy_data <- 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    values_to = "bdi")

pulse_tidy_data
```
Now `visit` is an issue. The original column names were informative but we probably don’t need to keep the `bdi_score_` prefix in each case. I’ll use an additional option in `pivot_longer` to address this.

```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi")

pulse_tidy_data
```
```{r}
pulse_df <- 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  relocate(visit) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)) %>%
  arrange(id, visit)

print(pulse_df, n = 12)
```
## `pivot_wider`

```{r}
analysis_result <- tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

analysis_result
```

```{r}
pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean")
```

## `bind_rows`

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6")%>%
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "gender", 
    values_to = "words") %>%
  mutate(race = str_to_lower(race)) %>% 
  select(movie, everything()) 
#never use rbind, always use bind_rows
lotr_tidy
```

